---
title: "Toy firms"
author: "Lars Vilhuber"
format: 
  html:
    keep-md: true

editor: source
---


A toy example of generating a somewhat plausible distribution of

- establishments across industries
- employment across establishments

Might still want to add payroll to approxmate the (Syn)LBD.

> A SAS version is available in directory [sas](sas/).

```{r setup,include=FALSE}
# Translation by Claude Sonnet 3.5
# Edits by Lars Vilhuber
library(knitr)
library(rprojroot)
basedir <- rprojroot::find_root(rprojroot::is_git_root)
indfig.png <- file.path(basedir,"ind-estab-counts.png")
sizefig.png <- file.path(basedir,"estab-size-dist.png")
```

## Basic parameters

```{r}
# Number of industries
indcnt <- 100
# Largest number of establishments per industry
maxestabcnt <- 10000
# Smallest industry
minestabcnt <- 100
seed1 <- 123456

# For any establishment, largest size
maxemp <- 42000
# smallest estab size
minemp <- 1
seed2 <- 1234567

```

Draw establishment count distribution across industries
- Use log normal distribution
- Set seeds for reproducibility

```{r}

set.seed(seed1)

# 
industries <- data.frame(
  industry = 1:indcnt,
  estabs = exp(runif(indcnt) * (log(maxestabcnt) - log(minestabcnt)) + log(minestabcnt))
)

```

# Summary statistics for estabs

```{r, results='asis'}
summary(industries$estabs) -> estabs_summary 
data.frame(
  Statistic = names(estabs_summary),
  Value = as.numeric(estabs_summary)) |>  kable()
```

```{r}
indfig <- hist(industries$estabs, main="Distribution of Establishments", xlab="Number of Establishments")
```


```{r,include=FALSE}
png(filename=indfig.png)
indfig
dev.off()
```


# Employment for each establishment in each industry

```{r}

set.seed(seed2)

fakelbd <- do.call(rbind, lapply(1:nrow(industries), function(i) {
  data.frame(
    industry = industries$industry[i],
    emp = exp(runif(industries$estabs[i]) * (log(maxemp) - log(minemp)) + log(minemp))
  )
}))

```

# Summary statistics for Employment (Establishment size)
```{r, results='asis'}
# For the employment summary
emp_summary <- summary(fakelbd$emp)
data.frame(
  Statistic = names(emp_summary),
  Value = as.numeric(emp_summary)
) |> kable(caption = "Summary of Employment")


```

```{r}
sizefig <- hist(fakelbd$emp, main="Distribution of Employment", xlab="Number of Employees")
```


```{r,include=FALSE}
png(filename=sizefig.png)
sizefig
dev.off()

```
